---
title: "Get Soil Properties"
author: "Ava Hoffman"
date: "2025-11-03"
---

```{r}
library(tidyverse) # %>% and other functions
library(sf)
```
---

The following come from Natural Resource Conservation Service, US Department of Agriculture. Gridded SSURGO (gSSURGO) is similar to the standard USDA-NRCS Soil Survey Geographic (SSURGO) Database product but in the format of an Environmental Systems Research Institute, Inc. (ESRI®) file geodatabase.

Landing page: https://www.nrcs.usda.gov/resources/data-and-reports/gridded-soil-survey-geographic-gssurgo-database 

Data can be downloaded by state here: https://nrcs.app.box.com/v/soils/folder/233398887779

There is a fair amount of work to determine which datasets and metadata to use. These have been helpful:

Data dictionary: https://www.nrcs.usda.gov/sites/default/files/2022-08/SSURGO-Metadata-Table-Column-Descriptions-Report.pdf
Column types and units: https://www.nrcs.usda.gov/sites/default/files/2022-08/SSURGO-Metadata-Tables-and-Columns-Report.pdf
Manually checking data: https://websoilsurvey.nrcs.usda.gov/app/WebSoilSurvey.aspx 

The following assumes the user has a directory at `../large_files/` containing the large files.

## Load data

```{r}
snapshot <- "BioDIGS_sites_20251111.csv"

full_data <-
  read.csv(
    paste0(
      "https://raw.githubusercontent.com/FredHutch/GDSCNsoilsites/refs/heads/main/data/snapshots/",
      snapshot
    )
  )
```

## Prepare data

```{r}
full_data <- tidyr::separate(full_data, gps, into = c("lat", "lon"), sep = ", ")
full_data$lat <- as.numeric(full_data$lat)
full_data$lon <- as.numeric(full_data$lon)
```

## Get state in the US

```{r}
# Data is quite large so this saves some time
full_data <- tidyr::separate(
  full_data,
  origin,
  into = c("city", "state"),
  sep = ", ",
  remove = FALSE
)
```

## Get soil survey data

```{r}
# Prepare coordinates DF
coords_df <- data.frame(
  state = full_data$state,
  site_name = full_data$site_name,
  lat = full_data$lat,
  lon = full_data$lon,
  Lat = full_data$lat,
  Lon = full_data$lon
)
```

```{r}
# Set up function to extract data
get_soil_survey <- function(state, coords_df) {
  # Contains polygons of soil map units (shapes)
  mapunits <- read_sf(paste0(
    "../large_files/gSSURGO_",
    state,
    ".gdb/a00000052.gdbtable"
  ))
  
  coords_df_state <- coords_df[coords_df$state == state, ]
  points_sf <- st_as_sf(coords_df_state,
                        coords = c("Lon", "Lat"),
                        crs = "+proj=longlat +datum=WGS84")
  points_sf <- st_transform(points_sf, st_crs(mapunits))
  points_sf_joined <- st_join(points_sf, mapunits, join = st_within) # takes a while!
  
  # Links map units to soil properties
  if (state %in% c("SC")) {
    # South Carolina's spatial tables are labeled a bit differently
    mapunits_components <- read_sf(paste0(
      "../large_files/gSSURGO_",
      state,
      ".gdb/a00000063.gdbtable"
    ))
  } else {
    mapunits_components <- read_sf(paste0(
      "../large_files/gSSURGO_",
      state,
      ".gdb/a00000061.gdbtable"
    ))
  }
  
  # Get the soil properties table
  soil_properties <- read_sf(paste0(
    "../large_files/gSSURGO_",
    state,
    ".gdb/a0000000f.gdbtable"
  ))
  
  points_sf_joined_soil <-
    points_sf_joined %>%
    left_join(mapunits_components, by = c("MUKEY" = "mukey")) %>%
    left_join(soil_properties, relationship = "many-to-many")
  
  soil_cleaned <- points_sf_joined_soil %>% select(
    state,
    site_name,
    lat,
    lon,
    AREASYMBOL,
    MUSYM,
    MUKEY,
    cokey,
    hzname,
    hzdept_r,
    hzdepb_r,
    sandtotal_r,
    silttotal_r,
    claytotal_r,
    om_r
  )
  
  return(soil_cleaned)
}
```

## Get Maryland Data

```{r}
soil_survey_out <- get_soil_survey("MD", coords_df)
full_data_soil_md <-
  dplyr::left_join(
    full_data,
    dplyr::distinct(soil_survey_out),
    by = c("lat", "lon", "state", "site_name")
  ) %>%
  filter(state == "MD") %>%
  filter(public_ok == TRUE)
readr::write_csv(full_data_soil_md, "data/gSSURGO_MD.csv")
```

## Get Virginia Data

```{r}
soil_survey_out <- get_soil_survey("VA", coords_df)
full_data_soil_va <-
  dplyr::left_join(
    full_data,
    dplyr::distinct(soil_survey_out),
    by = c("lat", "lon", "state", "site_name")
  ) %>%
  filter(state == "VA") %>%
  filter(public_ok == TRUE)
readr::write_csv(full_data_soil_va, "data/gSSURGO_VA.csv")
```

## Get North Carolina Data

```{r}
soil_survey_out <- get_soil_survey("NC", coords_df)
full_data_soil_nc <-
  dplyr::left_join(
    full_data,
    dplyr::distinct(soil_survey_out),
    by = c("lat", "lon", "state", "site_name")
  ) %>%
  filter(state == "NC") %>%
  filter(public_ok == TRUE)
readr::write_csv(full_data_soil_nc, "data/gSSURGO_NC.csv")
```

## Get South Carolina Data

```{r}
# There are actually some georgia sites listed under SC, so filter these out. They won't show up under SC's geometry.
coords_df_sc <-
  coords_df %>%
  filter(site_name != "Keysville, GA" &
           site_name != "Augusta, GA")
soil_survey_out <- get_soil_survey("SC", coords_df_sc)
full_data_soil_sc <-
  dplyr::left_join(
    full_data,
    dplyr::distinct(soil_survey_out),
    by = c("lat", "lon", "state", "site_name")
  ) %>%
  filter(state == "SC" &
           site_name != "Keysville, GA" &
           site_name != "Augusta, GA")  %>%
  filter(public_ok == TRUE)
readr::write_csv(full_data_soil_sc, "data/gSSURGO_SC.csv")
```

## Get Colorado Data

```{r}
soil_survey_out <- get_soil_survey("CO", coords_df)
full_data_soil_co <-
  dplyr::left_join(
    full_data,
    dplyr::distinct(soil_survey_out),
    by = c("lat", "lon", "state", "site_name")
  ) %>%
  filter(state == "CO") %>%
  filter(public_ok == TRUE)
readr::write_csv(full_data_soil_co, "data/gSSURGO_CO.csv")
```

## Get Texas Data

```{r}
soil_survey_out <- get_soil_survey("TX", coords_df)
full_data_soil_tx <-
  dplyr::left_join(
    full_data,
    dplyr::distinct(soil_survey_out),
    by = c("lat", "lon", "state", "site_name")
  ) %>%
  filter(state == "TX") %>%
  filter(public_ok == TRUE)
readr::write_csv(full_data_soil_tx, "data/gSSURGO_TX.csv")
```

## Get California Data

```{r}

```

## Get Georgia Data

```{r}
# There are actually some georgia sites listed under SC, so add these back. They won't show up otherwise.

```

## Get New York Data

```{r}

```

## Get New Jersey Data

```{r}

```

## Get Minnesota Data

```{r}

```

## Get Tennessee Data

```{r}

```

## Get North Dakota Data

```{r}

```

## Get Oklahoma Data

```{r}

```

## Get Washington Data

```{r}

```

## Get Arizona Data

```{r}

```

# Alternative data (FAO)

The following code helps find Soil Organic Carbon associated with each set of GPS coordinates.

Data are derived from the The Global Soil Organic Carbon Map v1.5 (GSOC), a product of GloSIS (Global Soil Information System) organization that is part of the Food and Agriculture Organization of the United Nations (FAO).

This data is from version 1.5, published in 2018.

Metadata/report can be found here: https://openknowledge.fao.org/server/api/core/bitstreams/c3ccec0d-fe75-49b7-9a4c-ee0a8777fed9/content

Data can be downloaded here: https://data.apps.fao.org/catalog//iso/7730e747-eb73-49c9-bfe6-84ebae718743 

Spatial scale is at 1km resolution and depth is 0-30cm. In the USA, SOC analysis method is primarily dry combustion and is derived from Natural Resource Conservation Service, US Department of Agriculture.

