---
title: "Get Soil Properties"
author: "Ava Hoffman"
date: "2025-11-03"
---

```{r}
library(tidyverse) # %>% and other functions
library(sf)
```
---

The following come from Natural Resource Conservation Service, US Department of Agriculture. Data can be downloaded by state here: https://nrcs.app.box.com/v/soils/folder/233398887779?page=2

There is a fair amount of work to determine which datasets and metadata to use. These have been helpful:

Data dictionary: https://www.nrcs.usda.gov/sites/default/files/2022-08/SSURGO-Metadata-Table-Column-Descriptions-Report.pdf
Column types and units: https://www.nrcs.usda.gov/sites/default/files/2022-08/SSURGO-Metadata-Tables-and-Columns-Report.pdf
Manually checking data: https://websoilsurvey.nrcs.usda.gov/app/WebSoilSurvey.aspx 

The following assumes the user has a directory at `../large_files/` containing the large files.

## Load data

```{r}
snapshot <- "BioDIGS_sites_20251103.csv"

full_data <-
  read.csv(
    paste0(
      "https://raw.githubusercontent.com/FredHutch/GDSCNsoilsites/refs/heads/main/data/snapshots/",
      snapshot
    )
  )
```

## Prepare data

```{r}
full_data <- tidyr::separate(full_data, gps, into = c("lat", "lon"), sep = ", ")
full_data$lat <- as.numeric(full_data$lat)
full_data$lon <- as.numeric(full_data$lon)
```

## Get state in the US

```{r}
# Data is quite large so this saves some time
full_data <- tidyr::separate(full_data, origin, into = c("city", "state"), sep = ", ", remove = FALSE)
```

## Get soil survey data

```{r}
# Prepare coordinates DF
coords_df <- data.frame(
  state = full_data$state,
  site_name = full_data$site_name,
  lat = full_data$lat,
  lon = full_data$lon,
  Lat = full_data$lat,
  Lon = full_data$lon
)
```

```{r}
# Set up function to extract data
get_soil_survey <- function(state, coords_df) {
  # Contains polygons of soil map units (shapes)
  mapunits <- read_sf(paste0(
    "../large_files/gSSURGO_",
    state,
    ".gdb/a00000052.gdbtable"
  ))
  
  coords_df_state <- coords_df[coords_df$state == state, ]
  points_sf <- st_as_sf(coords_df_state,
                        coords = c("Lon", "Lat"),
                        crs = "+proj=longlat +datum=WGS84")
  points_sf <- st_transform(points_sf, st_crs(mapunits))
  points_sf_joined <- st_join(points_sf, mapunits, join = st_within) # takes a while!
  
  # Links map units to soil properties
  mapunits_components <- read_sf(paste0(
    "../large_files/gSSURGO_",
    state,
    ".gdb/a00000061.gdbtable"
  ))
  
  # Get the soil properties table
  soil_properties <- read_sf(paste0(
    "../large_files/gSSURGO_",
    state,
    ".gdb/a0000000f.gdbtable"
  ))
  
  points_sf_joined_soil <- 
    points_sf_joined %>% 
    left_join(mapunits_components, by = c("MUKEY" = "mukey")) %>% 
    left_join(soil_properties, relationship = "many-to-many")
  
  soil_cleaned <- points_sf_joined_soil %>% select(
    state,
    site_name,
    lat,
    lon,
    AREASYMBOL,
    MUSYM,
    MUKEY,
    cokey,
    hzname,
    hzdept_r,
    hzdepb_r,
    sandtotal_r,
    silttotal_r,
    claytotal_r,
    om_r
  )
  
  return(soil_cleaned)
}
```

## Get Maryland Data

```{r}
soil_survey_out <- get_soil_survey("MD", coords_df)
full_data_soil_md <- dplyr::left_join(full_data, dplyr::distinct(soil_survey_out), by = c("lat", "lon", "state", "site_name")) %>% filter(state == "MD")
readr::write_csv(full_data_soil_md, "data/gSSURGO_MD.csv")
```

## Get Virginia Data

```{r}
soil_survey_out <- get_soil_survey("VA", coords_df)
full_data_soil_va <- dplyr::left_join(full_data, dplyr::distinct(soil_survey_out), by = c("lat", "lon", "state", "site_name")) %>% filter(state == "VA")
readr::write_csv(full_data_soil_va, "data/gSSURGO_VA.csv")
```

# Alternative data (FAO)

The following code helps find Soil Organic Carbon associated with each set of GPS coordinates.

Data are derived from the The Global Soil Organic Carbon Map v1.5 (GSOC), a product of GloSIS (Global Soil Information System) organization that is part of the Food and Agriculture Organization of the United Nations (FAO).

This data is from version 1.5, published in 2018.

Metadata/report can be found here: https://openknowledge.fao.org/server/api/core/bitstreams/c3ccec0d-fe75-49b7-9a4c-ee0a8777fed9/content

Data can be downloaded here: https://data.apps.fao.org/catalog//iso/7730e747-eb73-49c9-bfe6-84ebae718743 

Spatial scale is at 1km resolution and depth is 0-30cm. In the USA, SOC analysis method is primarily dry combustion and is derived from Natural Resource Conservation Service, US Department of Agriculture.

